---
title: "252_well_BC template"
output: html_document
editor_options: 
  chunk_output_type: console
---

The purpose of this script is to generate 2 csv files:
1. For Marvin: This should tell Marvin how much volume of each metal to pipette into each well of the final barcoding plates. Should contain essentially four fields specifying the following: origin plate, origin plate well, target plate, target plate well and volume. 

2. For debarcoding: This should specify which metals are present in each well for each final barcode plate. 

This template is for 252 well BC but can be modified to create other plae formats. 
Sections indicated with "!!!!" need to be input or adapted by the user.

#Load libraries
```{r}
library(gtools)
library(data.table)
```


#Create a data frame of metal combinations and randomise the order
```{r}

# !!!! Create a vector of all the metals you want to use
bc_metals <- c("102Pd", "103Rh", "104Pd", "105Pd", "106Pd", "108Pd", "110Pd", "139La") 
n <- length(bc_metals)

# !!!! Specify how many metals should be in each combination
k <- 4
combs <- combinations(n, k, v = bc_metals, set = TRUE, repeats.allowed = FALSE)

#Convert to a dataframe
combs <- as.data.frame(combs)

#Shuffle the row indices of the dataframe
set.seed(345)
rows <- sample(nrow(combs))

#Use this random vector to reorder the dataset
combs_rand <- combs[rows, ]
```


#Create a vector of the central 60 wells of the finished barcoding plate

```{r}
pcols <- c("B", "C", "D", "E", "F", "G")
prows <- sprintf("%02d", 2:11)
wells <- c()

for (i in pcols) {
  pwells <- paste0(i, prows)
  wells <- c(wells, pwells)
}
```


#Add the target plate and target wells to the list of randomised combinations
```{r}

#convert bc to a data table
bc <- as.data.table(combs_rand)

# !!!! vector of target plate names
plates <- c("plate_1", "plate_2")

# !!!! Add the target plate and wells per plate to the data table
bc[, plate := rep(plates, c(60, 10))]

#Add well names
bc <- bc[, .(V1, V2, V3, V4, plate, well = wells)]

# !!!! change column names
setnames(bc, c("metal_1", "metal_2", "metal_3", "metal_4", "plate", "well"))
```


#Generate the csv template for Marvin
```{r}
# !!!! Needs to be adapted depending on how many plates you have
bc_marvin <- melt(bc, 
     id.vars = c("plate", "well"),
     measure.vars = c("metal_1", "metal_2", "metal_3", "metal_4"),
     variable.name = "metal_index",
     value.name = "metal_id")


# !!!! Create a new data table containing info on how you arrange the metals in the origin plate
metal_key <- data.table(originpos = "P12", 
                        originwell = c("A02", "B02", "C02", "D02", 
                                       "E02", "F02", "G02", "H02"),
                        metal_id = bc_metals)

#Next join the data tables on the "metal_id" column
bc_marvin <- bc_marvin[metal_key, on = .(metal_id)]

# !!! Add a column to say which position on Marvin the targetplate will be located
bc_marvin[, targetpos := "P13"]

# !!!! Add a column for volume of each metal to be added per well
bc_marvin[, volume := 15]

#Change variable names
setnames(bc_marvin, c("targetplate", "targetwell", "metal_index", "metal_id", "originplate", "originwell", "targetpos", "volume"))

#Sort according to plate and well
setkey(bc_marvin, targetplate, targetwell)
bc_marvin

```



#Sanity checks
```{r}
#Does each metal occur the same number of times?
bc_marvin[, .N, by = metal_id]

```


#Create the key used for debarcoding
```{r}
#dcast the data table, assign a 0 where the metal is absent
bc_key <- dcast(bc_marvin, targetplate + targetwell ~ metal_id, value.var = "volume", fill = 0)

#Replace the volume value with 1 to indicate the metal is present
bc_key <- bc_key[, lapply(.SD, function(x) replace(x, which(is.numeric(x) == TRUE & x!=0), 1))]

# !!!! Change column and value names to be compatible with barcoding key in pipeline
setnames(bc_key, c("plate", "well", "Pd102", "Rh103", "Pd104", "Pd105", 
                   "Pd106", "Pd108", "Pd110", "La139"))

#Remove the prefix "plate_" from the plate variable values
bc_key$plate <- gsub("plate_", "", bc_key$plate)

bc_key
```


#Sanity checks
```{r}
#Does each plate have the correct number of wells?
bc_key[, .N, by = plate]

#Does each metal occur the right number of times?
bc_key[, .(sum(Pd102), sum(Rh103), sum(La139))]

```



#fwrite the csvs
```{r}
#Specify the file path to the folder where you want to save the csvs
bc_folder <- file.path("C:/Users/MichelleD/switchdrive/Islet_Project/Barcoding/202101_60WBC/")

#Specify file names for the csvs
fn.marvin <- paste0("bc_marvin", ".csv")
fn.key <-  paste0("bc_key", ".csv")
fn.metal_key <- paste0("metal_key_originplate", ".csv")

#Write the csvs
fwrite(bc_marvin, file.path(bc_folder, fn.marvin))
fwrite(bc_key, file.path(bc_folder, fn.key))
fwrite(metal_key, file.path(bc_folder, fn.metal_key))
```



